ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data

### 2023-11-10 manuscript/manuscript/manuscript.Rmd
Complete first draft of manuscript introduction

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data'
tags:
  - R
  - meta-analysis
  - metaresearch
  - ecology
  - evolutionary biology
  - many-analysts
authors:
  - name: Elliot Gould
    orcid: 0000-0003-0872-7098
    affiliation: "1" 
  - name: Hannah S. Fraser
    orcid: 0000-0000-0000-0000
    affiliation: "1"
  - name: Shinichi Nakagawa
    orcid: 0000-0000-0000-0000
    affiliation: "2"
  - name: Timothy H. Parker
    orcid: 0000-0000-0000-0000
    affiliation: "3"
affiliations:
 - name: The University of Melbourne
   index: 1
 - name: The University of New South Wales
   index: 2
 - name: Whitman College
   index: 3
citation_author: Gould and Fraser et al. 
date: "`r Sys.Date()`"
year: 2023
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

@@

*ManyEcoEvo* is an R package[@RCoreTeam2022] designed to *facilitate the analysis of* a many-analyst study in ecology and evolutionary biology, the *ManyEcoEvo Project* [@gould2023].
The purpose of *ManyEcoEvo* is to reproducibly implement the analysis and presentation of results for Gould *et al.* [-@gould2023] such that the 246 analysts as well as other interested researchers may independently reproduce the study's findings.
The package also aims to facilitate re-analyses and replications of Gould *et al.* [-@gould2023], as well as facilitating the analysis of similar many-analyst style datasets.

All raw data sets provided to independent teams for analysis, as well data collected on the analyses themselves are archived at **OSF Link1, OSF Link2,** respectively, and are provided in the GitHub repository for the package **(INSERT LINK)**, archived on Zenodo **(INSERT CITATION).** The ManyEcoEvo project contains tools for data-cleaning and processing tasks encountered when working with with many-analyst style data, functions for implementing and reproducing the analysis described inGould *et al.* [-@gould2023], establishes a structured workflow that facilitates re-analyses and replications of Gould *et al.* [-@gould2023] as well as implementation for new many-analyst studies, including a planned extension of the ManyEcoEvo project.

@@

The last ten years has seen the growing emergence of many-analyst projects, [e.g. @silberzahn2018; @botvinik-nezer2020; @huntington-klein2021; @schweinsberg2021; @breznau2022; @hoogeveen2022many; @coretta2023], predominantly in the disciplines of psychology and social and behavioural sciences, however we expect more of these types of studies to be published into the near future across other disciplines.
Moreover, collaborative and distributed big team open science projects are becoming increasingly common in ecology and evolutionary biology (e.g. NutNet <https://nutnet.org/index.php/>, DragNet, ManyBirds <http://themanybirds.com>) as well as other disciplines (ManyBabies **CITE**, ManyPrimates **CITE**) which pose computationally analogous data-manipulation and analysis problems to the many-analyst workflow addressed by the *ManyEcoEvo* approach.
While *ManyEcoEvo* was designed for the ManyEcoEvo project, the structured approach it employs establishes a workflow that can be generalised and applied to new many-analyst style data, facilitating replication of Gould *et al.* [-@gould2023], in addition to computational reproduction [@Stoudt2021], as well as new analyses of many-analyst type studies.

We are not aware of any R packages or software in other languages exclusively focussed on facilitating many-analyst style studies.
Existing many-analyst code if shared, is written as R scripts particular to that study's analysis (e.g. <https://github.com/manybabies/mb1-analysis-public>, <https://github.com/many-speech-analyses/many_analyses>).
However, there are a range of existing packages with select functionality that addresses various components of many-analyst style analyses, namely meta-analysis, its associated preparation and visualisation, as well as the visualisation component of [specification curve analysis] / [multiverse analysis] ([@Simonsohn2020], [@Steegen2016]).
Following the collection of results from teams conducting independent analyses, the data must be 1.
cleaned and prepared for 2.
meta-analysis and other model analysis, which is then 3.
subjected to model checking, statistic extraction and visualisation [@aczel2021].

@@

*ManyEcoEvo* provides custom functions to calculate standardize effect sizes in the form of Fishers' Z correlation scores in preparation for meta-analysis.
While some existing functions exist within other packages (e.g. `DescTools::FishersZ()`, `DiagTest3Grp::FisherZ.var()`), the method for calculating variance of the standardized effect size in Gould *et al.* [-@gould2023] differed slightly see `INSERT FUNCTION NAME`, and we wished to avoid reliance on too many external packages (Wickham & Bryan, 2023: **Chapter 10.1**).
Should *ManyEcoEvo* users wish to deviate from the effect-size standardisation method used in Gould *et al.* [-@gould2023], they are free to use their own functions provided by other software or custom functions (See **Harrer et al. 2022** chapter 3 for a range of R snippets that calculate various effect sizes).

`meta` (**CITE**) and `metafor` (**cite**) R packages are the two most commonly used softwares for undertaking meta-analysis in R [@lortie2020].
We provide functions for fitting the exact meta-analysis models used in Gould *et al.* [-@gould2023], which were implemented with the *metafor* package, however *ManyEcoEvo* users are free to supply alternative model-fitting functions should they wish to deviate from these specified models.
After model-fitting and analysis, model checking and extraction of model fit and heterogeneity statistics is implemented using a range of existing packages, including: `performance::`, `broom.mixed::`, `metafor::`.
*ManyEcoEvo* also draws on existing visualisation packages to implement forest plots (`ggforest::)`, funnel plots (`metafor::`), and s [specification curve analysis] plots to examine the influence of analysis specifications on variation in analysis outcomes (`specr::`).
Select functions from the above packages have been thoughtfully abstracted and coded into wrapper functions according to a principled and structured workflow (described below) for ease of application and generalisation to alternative datasets.

# Same Data, Different Analysts: variation in effect sizes due to analytical decisions in ecology and evolutionary biology.

>>>>

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data'
tags:
  - R
  - meta-analysis
  - metaresearch
  - ecology
  - evolutionary biology
  - many-analysts
authors:
  - name: Elliot Gould
    orcid: 0000-0003-0872-7098
    affiliation: "1" 
  - name: Hannah S. Fraser
    orcid: 0000-0000-0000-0000
    affiliation: "1"
  - name: Shinichi Nakagawa
    orcid: 0000-0000-0000-0000
    affiliation: "2"
  - name: Timothy H. Parker
    orcid: 0000-0000-0000-0000
    affiliation: "3"
affiliations:
 - name: The University of Melbourne
   index: 1
 - name: The University of New South Wales
   index: 2
 - name: Whitman College
   index: 3
citation_author: Gould and Fraser et al. 
date: "`r Sys.Date()`"
year: 2023
bibliography: paper.bib
output: 
  # word_document: default
  rticles::joss_article:
    journal: JOSS
csl: apa.csl
---

@@

*ManyEcoEvo* is an R package [@RCoreTeam2022] designed to *facilitate the analysis of* a many-analyst study in ecology and evolutionary biology, the *ManyEcoEvo Project* [@gould2023].
The purpose of *ManyEcoEvo* is to reproducibly implement the analysis and presentation of results for Gould *et al.* [-@gould2023] such that the 246 analysts as well as other interested researchers may independently reproduce the study's findings.
The package also aims to facilitate re-analyses and replications of Gould *et al.* [-@gould2023], as well as facilitating the analysis of similar many-analyst style datasets.

All raw data sets and metadata provided to independent teams for analysiss are publicly archived at <https://osf.io/qjzby> and <https://osf.io/hdv8m>, respectively.
Analyst survey responses and results are provided in the package GitHub repository (<https://github.com/egouldo/ManyEcoEvo/>), which is also archived on Zenodo (<https://doi.org/10.5281/zenodo.10046153>).
*ManyEcoEvo* contains tools for data-cleaning and processing tasks encountered when working with with many-analyst style data, functions for implementing and reproducing the analysis described inGould *et al.* [-@gould2023], establishes a structured workflow that facilitates re-analyses and replications of Gould *et al.* [-@gould2023] as well as implementation for new many-analyst studies, including a planned extension of the ManyEcoEvo project.

@@

The last ten years has seen the growing emergence of many-analyst projects, [e.g. @silberzahn2018; @botvinik-nezer2020; @huntington-klein2021; @schweinsberg2021; @breznau2022; @hoogeveen2022many; @coretta2023], predominantly in the disciplines of psychology and social and behavioural sciences, however we expect more of these types of studies to be published into the near future across other disciplines.
Moreover, collaborative and distributed big team open science projects are becoming increasingly common in ecology and evolutionary biology (e.g. NutNet/DragNet <https://nutnet.org/index.php/>, ManyBirds <http://themanybirds.com>) as well as other disciplines [ManyBabies @ManyBabies , ManyPrimates @ManyPrimates] which pose computationally analogous data-manipulation and analysis problems to the many-analyst workflow addressed by the *ManyEcoEvo* approach. While *ManyEcoEvo* was designed for the ManyEcoEvo project, the structured approach it employs establishes a workflow that can be generalised and applied to new many-analyst style data, facilitating replication of Gould *et al.* [-@gould2023], in addition to computational reproduction [@Stoudt2021], as well as new analyses of many-analyst type studies.

We are not aware of any R packages or software in other languages exclusively focussed on facilitating many-analyst style studies.
Existing many-analyst code if shared, is written as R scripts particular to that study's analysis (e.g. <https://github.com/manybabies/mb1-analysis-public>, <https://github.com/many-speech-analyses/many_analyses>).
However, there are a range of existing packages with select functionality that addresses various components of many-analyst style analyses, namely meta-analysis, its associated preparation and visualisation, as well as the visualisation component of specification curve analysis / multiverse analysis [@Simonsohn2020; @Steegen2016].
Following the collection of results from teams conducting independent analyses, the data must be 1.
cleaned and prepared for 2.
meta-analysis and other model analysis, which is then 3.
subjected to model checking, statistic extraction and visualisation [@aczel2021].

@@

*ManyEcoEvo* provides custom functions to calculate standardize effect sizes in the form of Fishers' Z correlation scores in preparation for meta-analysis.
While some existing functions exist within other packages [e.g.
`DescTools::FishersZ()` @DescTools, `DiagTest3Grp::FisherZ.var()` @luo2012diagtest3grp], the method for calculating variance of the standardized effect size in Gould *et al.* [-@gould2023] differed slightly, and we wished to avoid reliance on too many external packages [@wickham2023r, Chapter 10.1].
Should *ManyEcoEvo* users wish to deviate from the effect-size standardisation method used in Gould *et al.* [-@gould2023], they are free to use their own functions provided by other software or custom functions [@harrer2021, see chapter 3 for a range of R code to calculate various effect sizes].

`metafor` [@Viechtbauer:2010ks; @Viechtbauer2017] and `meta` [@Balduzzi2019] R packages are the two most commonly used softwares for undertaking meta-analysis in R [@lortie2020].
We provide functions for fitting the exact meta-analysis models used in Gould *et al.* [-@gould2023], which were implemented with the *metafor* package [@Viechtbauer2017], however *ManyEcoEvo* users are free to supply alternative model-fitting functions should they wish to deviate from these specified models.
After model-fitting and analysis, model checking and extraction of model fit and heterogeneity statistics is implemented using a range of existing packages, including: `performance::` [@ludecke2021], `broom.mixed::` [@Bolker2022], `metafor::` [@Viechtbauer2017].
*ManyEcoEvo* also draws on existing visualisation packages to implement forest plots [`ggforestplot::`, @ggforestplot; @orchaRd], funnel plots [`metafor::` @Viechtbauer2017], and specification curve analysis plots [`specr::` @Masur2020 ] to examine the influence of analysis specifications on variation in analysis outcomes.
Select functions from the above packages have been thoughtfully abstracted and coded into wrapper functions according to a principled and structured workflow (described below) for ease of application and generalisation to alternative datasets.

# Same Data, Different Analysts: variation in effect sizes due to analytical decisions in ecology and evolutionary biology

This is the body of the manuscript that will provide a short example of the claims about functionality I've made in the previous sections using the ManyEcoEvo analysis pipeline for correlation coefficients.
